### 2020-04-10 paper/main.tex
typo
new updates

   By Ruben Rosales \\
   May 2020
\end{center}

@@

In this thesis, we study interpretable machine learning as applied to complex-valued time-series data. Scientists have studied the use of several machine learning methods such as Convolutional Neural Networks, Recurrent Neural Networks, and Support Vector Machines for time series classification. These methods, however, fall short of allowing users to visualize patterns within their dataset.

@@

\tableofcontents
\listofillustrations

@@

% \section{The need for interpretable models in management problems}
% The growth of mobile devices and demand for wireless data has created a need for igh quality spectrum sensing and adaptation to improve spectral allocation and interference mitigation is an important route by which we may achieve this.  however have been constrained to relatively specialized solutions which lack the generality needed to deal with a complex and growing number emitter types, interference types and propagation environments.
% This is a significant challenge in the community as expert systems designed to perform well on specialized tasks often lack flexibility and can be expensive and tedious to develop analytically.

\section{Contributions}
We focus on creating deep learning architectures to classify our datasets as well as extending existing work to create our white-box model. Our final results show how robust our black-box model is with its competitive accuracy. Additionally, we show the effectiveness of our white-box model in understanding the features of a dataset not normally exposed through a black-box model.

@@

We analyzed our model on three radio signal datasets, which we will refer to as dataset A, B, and C throughout this thesis. These datasets were generated through Spatial Modulation which is a transmission technique that uses multiple antennae which maps a block of information bits to two units, a symbol chosen from a constellation diagram, and a unique transmit antenna number that is chosen from a set of transmit antennae \cite{mesleh_spatial_2008}. The method in which transmit antennae send and receive radio waves can be described through polarization. Two popular basis polarizations are horizontal linear, H, and vertical linear, V \cite{oshea_convolutional_2016}. Our datasets contain data horizontally transmitted and vertically received (HV), as well as data vertically transmitted and vertically received (VV).

@@

\begin{figure}[h!]
   \begin{center}
      \includegraphics[scale=0.4]{../media/dataset_ac.png}
   \end{center}
   \caption{Class representation of how dataset A and C.}
   \label{fig:dataset_ac}
\end{figure}

@@

\begin{figure}[h!]
   \begin{center}
      \includegraphics[scale=0.5]{../media/dataset_b.png}
   \end{center}
   \caption{Class representation of how dataset B.}
   \label{fig:dataset_b}
\end{figure}

@@

\chapter{Related Work}
\label{chap:relatedwork}

The study of interpretable hybrid models has been worked on extensively in the past few years as deep learning has become increasingly popular. Scientists take different approaches as to what makes a hybrid model interpretable and apply it to different domains, such as computer vision or time series. For example, in \cite{wang_multilevel_2018}, they focus on creating a black-box model for the classification of time-series that outperforms similar architectures. However, their white-box model focuses on analyzing the importance of their model's layers. This helps users identify the importance of wavelet decomposition but not necessarily understand their dataset.

In order to fill the gap for a high performing interpretable model for time-series classification, we propose our hybrid model, which focuses on extensibility and robustness. We achieve this through a combination of deep learning models for classification as well as a unique white-box method to explaining features within classes of a given dataset.

@@

% [fft formula]
The Fourier Transform (FT) is a mathematical tool that decomposes any function into a sum of sinusoidal basis functions. Each basis function is a complex value of a frequency, allowing us to view our data in the frequency domain as opposed to the amplitude domain.

@@

@@

\begin{figure}[h!]
   \begin{center}
      \includegraphics[scale=0.4]{../media/wavelet_families.png}
   \end{center}
   \caption{Example of different types of wavelets \cite{lopez_development_2017}.}
   \label{fig:reinforcementAgent}
\end{figure}

@@

\subsection{Each Level of Decomposition as a Dataset}
Figure \ref{fig:multimodal_overview} depicts our model in which each level of decomposition is a unique dataset. This method and the resampling method both achieved similar results, so we chose to move forward with this method.

@@

\section{Definition}
Machine learning algorithms can be seen as learning a target function (f) that maps input data (X) to an output (Y). There are several techniques to make this work, but we focus on nonparametric algorithms, namely Convolutional Neural Networks (CNN) and Long Short-Term Memory networks (LSTM). Nonparametric algorithms attempt to make minimal assumptions about the form of the function to learn any form from data provided to it. An example would be a neural network as it has no prior knowledge of what it is classifying and attempts to generalize any new data points.

@@

A Convolutional Neural Network is a type of deep neural network that can be applied to different domains such as computer vision or time series analysis. As seen in Figure X, it consists of an input and output layer as well as several hidden layers. A convolution is an operation between a vector of weights w against an input x \cite{yamashita_convolutional_2018}. It consists of taking the dot product between m and x in steps of a filter size defined by n.

@@

\begin{figure}[h!]
   \begin{center}
      \includegraphics[scale=0.6]{../media/results.png}
   \end{center}
   \caption{Results}
   \label{fig:final_results}
\end{figure}

To test our model, we split our data into training and test sets by randomly selecting 70\% of our data as our training set and leaving the other 30\% to be our test set. We then performed three-fold cross-validation and took the average of our results as the outcome.

@@

We then split those images using a randomized 70/30 split where 70\% of the data is our training set, and the other 30\% is our test set. We were able to achieve 60\% accuracy across all three of our datasets and show the features amongst the classes behave differently, as can be seen in Figure \ref{fig:whitebox_images}.

@@

\begin{appendices}
   \chapter{MTF Code}
      \section{Main File}
         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible ]{../code/mtf/main.py}
      \section{Calculating MTF}
         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/mtf/complex_signal.py}
      \section{Helper functions}
         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/mtf/util.py}
   \chapter{Hybrid Model Code}
      \section{Main File}
         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/hybrid_model/main.py}
      \section{White-Box Model}
         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/hybrid_model/white_box.py}
      \section{Black-Box Model}
         \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/hybrid_model/black_box.py}

\end{appendices}

>>>>

   By Ruben Rosales \\
   B.S., 2016, California State University, Fullerton \\
   May 2020
\end{center}

@@

In this thesis, we study interpretable machine learning as applied to complex-valued time-series data. Scientists have studied the use of several machine learning methods such as Convolutional Neural Networks, Recurrent Neural Networks, and Support Vector Machines for time series classification. These methods, however, fall short of allowing users to visualize patterns, which can be a necessity for adoption in highly scrutinized industries.

@@

\tableofcontents
\listoffigures

@@

% \section{The need for interpretable models in management problems}
% The growth of mobile devices and demand for wireless data has created a need for igh quality spectrum sensing and adaptation to improve spectral allocation and interference mitigation is an important route by which we may achieve this.  however have been constrained to relatively specialized solutions which lack the generality needed to deal with a complex and growing number emitter types, interference types and propagation environments.
% This is a significant challenge in the community as expert systems designed to perform well on specialized tasks often lack flexibility and can be expensive and tedious to develop analytically.
\chapter{Related Work}
\label{chap:relatedwork}

The study of interpretable hybrid models has been worked on extensively in the past few years as deep learning has become increasingly popular. Scientists take different approaches as to what makes a hybrid model interpretable and apply it to different domains, such as computer vision or time series. For example, Wang \cite{wang_multilevel_2018}, focuses on creating a black-box model that utilizes wavelet decomposition for the classification of time-series that outperforms similar architectures. However, their white-box model focuses on analyzing the importance of their model's layers. This helps users identify the importance of wavelet decomposition but not necessarily understand their dataset.

In order to fill the gap for a high performing interpretable model for time-series classification, we propose our hybrid model, which focuses on extensibility and robustness. We achieve this through a combination of deep learning models for classification as well as a unique white-box method to explain features within classes of a given dataset.

% \section{Contributions}
% We focus on creating deep learning architectures to classify our datasets as well as extending existing work to create our white-box model. Our final results show how robust our black-box model is with its competitive accuracy. Additionally, we show the effectiveness of our white-box model in understanding the features of a dataset not normally exposed through a black-box model.

@@

We analyzed our model on three radio signal datasets, which we will refer to as dataset A, B, and C throughout this thesis. These datasets were generated through Spatial Modulation, which is a transmission technique that uses multiple antennae to map a block of information bits to two units, a symbol chosen from a constellation diagram, and a unique transmit antenna number that is chosen from a set of transmit antennae \cite{mesleh_spatial_2008}. The method in which transmit antennae send and receive radio waves can be described through polarization. Two popular basis polarizations are horizontal linear, H, and vertical linear, V \cite{oshea_convolutional_2016}. Our datasets contain data horizontally transmitted and vertically received (HV), as well as data vertically transmitted and vertically received (VV).

@@

\begin{figure}[h!]
   \begin{center}
      \includegraphics[scale=0.4]{../media/dataset_ac.png}
   \end{center}
   \caption{Class representation of dataset A and C.}
   \label{fig:dataset_ac}
\end{figure}

@@

\begin{figure}[h!]
   \begin{center}
      \includegraphics[scale=0.5]{../media/dataset_b.png}
   \end{center}
   \caption{Class representation of dataset B.}
   \label{fig:dataset_b}
\end{figure}

@@

@@

% [fft formula]
The Fourier Transform (FT) is a mathematical tool that decomposes any function into a sum of sinusoidal basis functions. Each basis function is a complex value of a frequency, allowing us to view our data in the frequency domain as opposed to the amplitude domain. Our dataset are discrete in time domain so we focus on discrete fourier transforms. The discrete fourier transform of a signal x can be defined as

\begin{equation} \label{eq1}
   \begin{split}
      X_{k} =  \sum_{n=0}^{N-1} x(n) e^{-\frac{i2\pi kn}{N}}
   \end{split}
\end{equation}

where x(t\textsubscript{n}) is the input signal at time n, X(k) is the \textit{k}th spectral sample \cite{smith_mathematics_nodate}.

@@

We can explain STFT through an example. Let x(n) be a signal, and let X\textsubscript{n}(e\textsuperscript{j$\omega\kappa$}) be the short-time Fourier transform of x(n) at time n and frequency $\omega$\textsubscript{$\kappa$} (Figure \ref{fig:stft_example}) \cite{allen_unified_1977}.

\begin{figure}[h]
   \begin{center}
      \includegraphics[scale=0.6]{../media/stft_eq.png}
   \end{center}
   \caption{Example of short-time fourier transform formula \cite{allen_unified_1977}.}
   \label{fig:stft_example}
\end{figure}

@@

% \begin{figure}[h!]
%    \begin{center}
%       \includegraphics[scale=0.4]{../media/wavelet_families.png}
%    \end{center}
%    \caption{Example of different types of wavelets \cite{lopez_development_2017}.}
%    \label{fig:reinforcementAgent}
% \end{figure}

@@

\subsection{Each Level of Decomposition as a Dataset}
Figure \ref{fig:wavelet_decomp} depicts our model in which each level of decomposition is a unique dataset. This method and the resampling method both achieved similar results, so we chose to move forward with this method.

@@

\section{Definition}
Machine learning algorithms can be seen as learning a target function (f) that maps input data (X) to an output (Y). Then, f is used to make predictions on new input data. There are several techniques to make this work, but we focus on nonparametric algorithms, namely Convolutional Neural Networks (CNN) and Long Short-Term Memory networks (LSTM). Nonparametric algorithms attempt to make minimal assumptions about the form of the function to learn any form from data provided to it. An example would be a neural network as it has no prior knowledge of what it is classifying and attempts to generalize any new data points.

@@

A Convolutional Neural Network is a type of deep neural network that can be applied to different domains such as computer vision or time series analysis. As seen in Figure \ref{fig:cnn_example}, it consists of an input and output layer as well as several hidden layers. A convolution is an operation between a vector of weights w against an input x \cite{yamashita_convolutional_2018}. It consists of taking the dot product between m and x in steps of a filter size defined by n.

@@

To test our model, we split our data into training and test sets by randomly selecting 70\% of our data as our training set and leaving the other 30\% to be our test set. We then performed three-fold cross-validation and took the average of our results as the outcome. As can be seen in Figure \ref{fig:final_results}, dataset A and C achieved similar accuracy, around 99\%, and dataset B averaged at around 91\% accuracy.

\begin{figure}[h!]
   \begin{center}
      \includegraphics[scale=0.6]{../media/results.png}
   \end{center}
   \caption{Accuracy on test data for our black-box model.}
   \label{fig:final_results}
\end{figure}

@@

We then split those images using a randomized 70/30 split where 70\% of the data is our training set, and the other 30\% is our test set. Afterward, we trained our CNN (Figure \ref{fig:mtf_cnn_model}) on those images and were able to achieve 60\% accuracy across all three of our datasets and show the features amongst the classes behave differently, as can be seen in Figure \ref{fig:whitebox_images}.

@@

% \begin{appendices}
%    \chapter{MTF Code}
%       \section{Main File}
%          \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible ]{../code/mtf/main.py}
%       \section{Calculating MTF}
%          \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/mtf/complex_signal.py}
%       \section{Helper functions}
%          \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/mtf/util.py}
%    \chapter{Hybrid Model Code}
%       \section{Main File}
%          \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/hybrid_model/main.py}
%       \section{White-Box Model}
%          \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/hybrid_model/white_box.py}
%       \section{Black-Box Model}
%          \lstinputlisting[language=Python,  basicstyle=\small, columns=fullflexible]{../code/hybrid_model/black_box.py}

% \end{appendices}
